Sherlock topic model paper

### 2018-05-23 main.tex
saving progress

\subsubsection{Measuring the quality of recall events}
Representing the video and verbal recall as events in `topic space' affords us the ability to characterize the quality of recall in a more fine-grained and nuanced way than what is typically possible. To quantify the similarity between the video model and individual recall models, we considered a number of novel metrics.  First, we tested whether each participant's recall model matched the movie model in a general sense. To do this, for each participant we filtered the video model to only include the events that the participant remembered. Then, we computed the root mean squared difference (RMSD) between the video model and the recall model. As an example, if the participant remembered all the events in order (with perfect precision), the expected distance value would be 0. However, if they remembered a subset of events, events our of order or with low precision the expected distance would be greater than 0. To assess significance, we performed a permutation test where we circularly shifted the recall model (10000 times) and recomputed the RMSD. The recall model significantly matched the video model for nine of the participants (p<.05; participants: 3-4, 8-13, 17 and the p-value for the rest of the participants was less than .25). Furthermore, the RMSD values were significantly correlated to memory performance across participants (Pearson's r=-.57, p=.016). Thus, a closer match between the video and recall event models was related to better recall performance.

Next, we tested whether participants who recalled more events were also more precise in their recollections. For each participant, we computed the correlation between each recall event and its matching video event (only for the events which they recalled). This resulted in a single number for each recalled event indexing how similar the recall event was to its matching movie event (i.e the `precision' of the recall). We then averaged the correlations within participant. In line with our prediction, there was a strong correlation between memory performance and precision suggesting that participants who remembered more events also remembered them more veridically (Pearson's r=.74, p=.0006). Next, we considered the distinctiveness of each recall event. That is, how uniquely a recall event matched a given video event compared to all other video events. We hypothesized that participants with high memory performance might describe each event in a more distinctive way (relative to those with lower memory performance who might describe events in a more general way). To this end, we computed a `distinctiveness' score for each participant (i.e. 1 - the correlation between a recall event and all non-matching video events).  Then, we averaged this measure over recall events within participant.  We found that participants with higher memory performance also had higher distinctiveness scores (Pearson's r=.8, p=.0001). Lastly, we tested whether participants with better memory performance were also more likely to remember the events in order.  For each participant, we computed the Spearman rank correlation between the order of events that the participant recalled and the actual order of events (filtering events that were actually recalled).  We found that participants who recalled more events also recalled more of them in order (Pearson's r=.5, p=.04). In summary, we found that better memory performance was associated with more precise and distinctive recall.

\begin{figure}[t!]
\centering
\includegraphics[width=1\textwidth]{figs/3_behav_quality.pdf}
\caption{\label{fig:behav}
}
\end{figure}

\subsubsection{Naturalistic extensions of classic memory analyses}
In addition to the novel metrics we introduce above, our approach also allows us to extend classic memory analyses designed to characterize memory in list-learning experiments to more naturalistic settings. Just like in a `free-recall' list-learning experiment where participants verbally recall a series of unique items in some order, our video-recall matching approach allows to analyze memory in the same way.

@@

\section{Methods}
\label{sec:methods}

>>>>

\subsubsection{Measuring the quality of recall events}
Representing the video and verbal recall as events in `topic space' affords us the ability to characterize the quality of recall in a more fine-grained and nuanced way than what is typically possible. To quantify the similarity between the video model and individual recall models, we considered a number of novel metrics.  First, we tested whether each participant's recall model matched the movie model in a general sense. To do this, for each participant we filtered the video model to only include the events that the participant remembered. Then, we computed the root mean squared difference (RMSD) between the video model and the recall model. As an example, if the participant remembered all the events in order (with perfect precision), the expected distance value would be 0. However, if they remembered a subset of events, events our of order or with low precision the expected distance would be greater than 0. To assess significance, we performed a permutation test where we circularly shifted the recall model (10000 times) and recomputed the RMSD. The recall model significantly matched the video model for nine of the participants [p<.05; participants: 3-4, 8-13, 17 and the p-value for the rest of the participants was less than .25]. Furthermore, the RMSD values were significantly correlated to memory performance across participants [Pearson's r(16)=-.57, p=.016]. Thus, a closer match between the video and recall event models was related to better recall performance.

% \begin{figure}[t!]
% \centering
% \includegraphics[width=1\textwidth]{figs/3_behav_quality.pdf}
% \caption{\label{fig:behav}
% }
% \end{figure}

Next, we tested whether participants who recalled more events were also more precise in their recollections. For each participant, we computed the correlation between each recall event and its matching video event (only for the events which they recalled). This resulted in a single number for each recalled event indexing how similar the recall event was to its matching movie event (i.e the `precision' of the recall). We then averaged the correlations within participant. In line with our prediction, there was a strong correlation between memory performance and precision suggesting that participants who remembered more events also remembered them more veridically [Pearson's r(16)=.74, p=.0006]. Next, we considered the distinctiveness of each recall event. That is, how uniquely a recall event matched a given video event compared to all other video events. We hypothesized that participants with high memory performance might describe each event in a more distinctive way (relative to those with lower memory performance who might describe events in a more general way). To this end, we computed a `distinctiveness' score for each participant (i.e. 1 - the correlation between a recall event and all non-matching video events).  Then, we averaged this measure over recall events within participant.  We found that participants with higher memory performance also had higher distinctiveness scores [Pearson's r(16)=.8, p=.0001]. Lastly, we tested whether participants with better memory performance were also more likely to remember the events in order.  For each participant, we computed the Spearman rank correlation between the order of events that the participant recalled and the actual order of events (filtering events that were actually recalled).  We found that participants who recalled more events also recalled more of them in order [Pearson's r(16)=.5, p=.04]. In summary, we found that better memory performance was associated with more precise and distinctive recall.

\subsubsection{Naturalistic extensions of classic memory analyses}
In addition to the novel metrics we introduce above, our approach also allows us to extend classic analyses designed to characterize memory in list-learning experiments to more naturalistic settings. Just like in a `free-recall' list-learning experiment where participants view a set of items (such as words) and then verbally recall them, our video-recall matching approach allows us to analyze memory in the same way (where the items are recall events). Since each recall event has a matching video event, we can consider characteristics of the sequence of recalls.

First, we asked whether the estimated number of recall events (k) by participant was related to hand-annotated accuracy as published in Chen et al. (2017).  We found a strong positive correlation where subjects with a greater number of recall events also had better memory performance (Fig~\ref{eventseg_behav}a; Pearson's $r$(16)=.67, $p$=.003).  Thus, to the extent that they are correlated, k could be use in place of a measure such as overall recall accuracy. We considered two additional across-participant measures of recall that characterize memory organization: temporal clustering and semantic clustering. Temporal clustering (as described in REF) measures the extent to which participants tend to cluster recalls of events encoded nearby in time.  For example, if a participant recalled in exact serial order, the score would be 1 whereas if they sequentially recalled randomly positioned items the score would be ~.5.  We found that participants who clustered in time also recalled a greater number of events (Fig~\ref{eventseg_behav}b; Pearson's $r$(16)=.62, $p$=.007). Next, we assessed semantic clustering.  Similar to temporal clustering, semantic clustering measures the extent to which participants cluster their responses by semantic similarity.  We used the topic vector representations of each event as a proxy for semantic information contained in that event.  For each transition, we computed the rank percentile of the of the correlation between the the current event and then next event. For example, if a participant sequentially recalled events where each successive recall was the event that was most semantically similar to the previous, the semantic clustering score would be 1.  If the events were recalled in a random order with respect to semantic information, the clustering score would be ~.5. We found that the semantic clustering score was related to memory performance across participants (Fig~\ref{eventseg_behav}c; Pearson's $r$(16)=.55, $p$=.02).  Thus, participants who organized their recalls with respect to the semantic information contained in the scene had better memory performance.

Next, we considered initiating the recall sequence (known in the literature as the `probability of first recall' or `PFR'). This metric quantifies the likelihood of initiating a recall sequence based on the item's serial position. We found that participants tended to initiate their recall sequences with the first few events, which is qualitatively very similar to previously published list learning experiments (REF).  Next, we considered another well-studied memory measure in list-learning, the lag conditional response probability curve (or lag-CRP). Given the recall of a stimulus in study position $n$, the lag-CRP quantifies the probability that the next item recalled will be of lag $i$. The results suggest a strong bias to transition sequentially in the forward direction. This can be seen by the high probability values in the lag ~1-5 positions compared to the rest of the transition possibilties. Notably, participants were instructed to recall as much as they could in order, and the shape of this lag-CRP is qualitatively similar to list-learning studies of serial recall (REF), where participants are required to recall lists of words in their serial order. Finally, we assessed memory performance for each event in the video. This plot quantifies the proportion of participants that recalled each event in the video (Fig~\ref{eventseg_behav}f).

@@

\subsubsection{Duration of recall events}

% \section{Methods}
% \label{sec:methods}
